---
title: "Lab5 Team"
author: "Jian Kai Lee;jiankailee / Nick Matthews;nickmatthews713"
date: "4/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Loading in files  
```{r warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(maps)
library(mapproj)
library(tidyr)
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
person <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)

```

### Part 1: Accident Data

#### Problem 1
```{r message=FALSE}
acc %>%ggplot(aes(x=DAY_WEEK))+geom_histogram()+xlab("Day of Week")+ylab("number of accidents") +
  ggtitle("Accidents Per Day of the Week")
```

##### Description:  
Accidents occurred more often in the first, six and seven day of week which are Monday, Saturday and Sunday. As we can see from the graph, the second, third and fourth are less accident than on the weekend.

#### Problem 2
```{r warning=FALSE, message=FALSE}
acc%>%ggplot(aes(x=HOUR))+geom_histogram()+xlim(0,24)
```

##### Description:  
From the histogram, we can see the most accidents happened between 17 and 21 which are 5 pm to 9pm. People tend to drive home from work or go out for dinner.

#### Problem 3
```{r}
temp<-acc%>%filter(DRUNK_DR>=1)
length(temp)
```

##### Description:  
The number of accidents with at least one drunk driver are 52.

### part 2: Connecting Data  
#### Problem 4
```{r}
person_sub<-person%>%filter(PER_TYP==1)
```

##### Description:  
when "PER_TYP" = 1, it refers to the person driving the motor vehicle

#### Problem 5
```{r}
joined <- inner_join(person_sub, acc, by=c("ST_CASE","DAY","MONTH","HOUR","MINUTE","STATE","VE_FORMS","COUNTY"))
```

##### Description:  
I assume we are trying to join together the data and try to match each person to the accident, so I made sure to join along many details down to the minute of when the crash happened.

#### Problem 6
```{r}
joined$SEX <- as.factor(joined$SEX)
levels(joined$SEX) <- c("Male","Female","Not Reported","Unknown")

joined$DAY_WEEK <- as.factor(joined$DAY_WEEK)
levels(joined$DAY_WEEK) <- c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday")

joined %>%
  group_by(DAY_WEEK, HOUR, SEX) %>%
  summarise(count = n()) %>%
  filter(HOUR < 25) %>%
  ggplot(aes(x=HOUR, y=count, fill=SEX)) + geom_bar(stat="identity") + facet_wrap(~DAY_WEEK) +
  ylab("Number of Crashes") + xlab("Hour of Day") + ggtitle("Number of Crashes for Each Day of the Week")
```

##### Description:  
Here is the amount of accidents for every hour for each day of the week, also displaying the gender by color. I factored the columns SEX and DAY_WEEK to make the graph more readable. I had to do a filter(HOUR < 25) because there is a row in the dataset where the HOUR is 100. This didn't make any sense, so I removed it.

### Part 3: Connecting Data and Make Maps  
#### Problem 7/8  
```{r warning=FALSE}
countys = map_data('county')
states = map_data('state')
GLC = readxl::read_xlsx('counties.xlsx')
GLC$`County Name` = tolower(GLC$`County Name`)
GLC$`State Name` = tolower(GLC$`State Name`)

countyCodes <- unique(GLC %>% select(`County Name`,`County Code`))

# summarise the accident date to get number of crashes per county
countyCases <- acc %>%
  group_by(COUNTY) %>% summarise(crashes = n())

# join together data to geocode and then use the polynames to get longitude and latitude from counties
countyData <- countyCases %>%
  inner_join(countyCodes %>% mutate('County Code' = as.numeric(countyCodes$'County Code')), 
             by=c(COUNTY='County Code'))  %>%
  inner_join(countys, by=c(`County Name` = 'subregion'))

# plot the states map along with all the counties on it, coloring the counties according to the number of crashes
countys %>%
  ggplot(aes(x=long, y=lat)) + geom_path(aes(group = group)) + coord_map() +
  geom_polygon(data = countyData, aes(x=long, y=lat, group=group, fill=crashes),
             size=0.05) +
  xlim(c(-125,-65)) + ylim(c(20,50))
```

#### Problem 9
```{r}
summerAcc <- acc %>% filter(MONTH < 4 | MONTH > 11) # Winter
winterAcc <- acc %>% filter(MONTH < 9 | MONTH > 4) #Summer

GLC$'State Code' = as.numeric(GLC$'State Code')
stateCodes <- unique(GLC %>% select(`State Name`,`State Code`))

winterAcc %>%
  group_by(STATE) %>%
  summarise(crashes = n()) %>%
  inner_join(stateCodes, by=c(STATE="State Code")) %>%
  inner_join(states, by=c(`State Name`="region")) %>%
  ggplot(aes(x=long,y=lat)) +
  geom_polygon(aes(group=group, fill = crashes)) +
  ggtitle("Crashes by State During the Winter") +
  scale_fill_gradientn(limits=c(0,3500), 
                       colours=c("navyblue", "darkmagenta", "darkorange1"))

summerAcc %>%
  group_by(STATE) %>%
  summarise(crashes = n()) %>%
  inner_join(stateCodes, by=c(STATE="State Code")) %>%
  inner_join(states, by=c(`State Name`="region")) %>%
  ggplot(aes(x=long,y=lat)) +
  geom_polygon(aes(group=group, fill = crashes)) +
  ggtitle("Crashes by State During the Summer") +
  scale_fill_gradientn(limits=c(0,3500), 
                       colours=c("navyblue", "darkmagenta", "darkorange1"))
  
```

##### Description:  
It appears that there are more crashes in just about every state during the winter(Especially for the midwest). This is because as the temp gets colder, roads freeze and cars slip easier, leading to more crashes.
